---
title: "Figure 1C"
format: html
theme: 
  light: zephyr
  dark: cyborg
toc: true
page-layout: full
number-sections: true
number-depth: 3
editor: visual
self-contained: true
---

# Figure 1C - Minmax normalized values

This Quarto document contains the data manipulation and plot generation for Figure 1C of Quijano, et al., 2024.

## Import libraries

```{r}
pacman::p_load(tidyverse,
               broom,
               ggbeeswarm,
               cowplot,
               tidytext,
               ggtext)
```

## Import metadata

```{r}
#| warning: false
A_metadata <- 
  read_csv('../data/Fig01_metadata.csv')
```

## Make a colors list

```{r}
Z_color_htol_int <-
  c("Thermotolerant"="#B2182B",
    "Intermediate"="gray",
    "Thermosensitive"="#2166AC")
```

## Import mortality data

```{r}
B_mor_df <- 
  read_csv('../data/Figure01C_Merged_Mortality.csv')

B_mor_df <- 
  B_mor_df |> 
  select(Species, Colony, `Proportion of fragments bleached, %`)
```

## Import MIG data

```{r}
B_mig_df <- 
  read_csv('../data/Figure01C_Merged_MIG.csv')

B_mig_df <- 
  B_mig_df |> 
  select(`Colony ID`, Mean) |> 
  rename(Colony = `Colony ID`)
```

## Normalize data

Merge.

```{r}
C_mor.mig_df <- 
  B_mor_df |> 
  left_join(B_mig_df,
            by = 'Colony')
```

Make a function for rescaling (min-max normalizaiton).

```{r}
Z_mor.mig_minmax_f <- 
  function(x){
    (x - min(x)) /(max(x) - min(x))
  }
```

Invert, normalize and combine data.

```{r}
C_mor.mig_df <- 
  C_mor.mig_df |> 
  group_by(Species) |> 
  
  # Invert values
  mutate(mor_norm = Z_mor.mig_minmax_f(`Proportion of fragments bleached, %`),
         mig_norm = Z_mor.mig_minmax_f(Mean)) |> 
  mutate(combined = (mor_norm + mig_norm))
```

Rank.

```{r}
C_mor.mig_rank <- 
  C_mor.mig_df |> 
  group_by(Species) |> 
  arrange(combined, .by_group = TRUE) |> 
  mutate(Ranking  = seq_along(Species)) |> 
  mutate('Heat tolerance' = case_when(Species == 'A. digitifera' & Ranking <= 8 ~ 'High',
                                      Species == 'A. digitifera' & Ranking >= 33 ~ 'Low',
                                      Species == 'A. millepora' & Ranking <= 6 ~ 'High',
                                      Species == 'A. millepora' & Ranking >= 25 ~ 'Low',
                                      Species == 'A. tenuis' & Ranking <= 5 ~ 'High',
                                      Species == 'A. tenuis' & Ranking >= 21 ~ 'Low',
                                      .default = 'Intermediate'))
```

Export.

```{r}
write_csv(C_mor.mig_rank,
          '../Results/Tables/Supplementary table 06 - minmax normalized values.csv')
```

Plot.

```{r}
C_mor.mig_p <- 
  C_mor.mig_rank |> 
  mutate(`Heat tolerance` = case_when(`Heat tolerance` == "High" ~ "Thermotolerant",
                                      `Heat tolerance` == "Low" ~ "Thermosensitive",
                                      .default = `Heat tolerance`),
         Species = if_else(Species == "A. tenuis", 
                           "<i>A. </i>cf. <i>tenuis</i>", 
                           paste0("<i>",Species,"</i>"))) |> 
  ggplot() +
  geom_bar(aes(x = reorder(`Colony`, -combined),
               y = mor_norm),
           stat = 'identity',
           alpha = 0.15) +
  geom_point(aes(x = reorder(`Colony`, -combined),
               y = combined,
               fill = fct_relevel(`Heat tolerance`, c("Thermosensitive",
                                                      "Intermediate",
                                                      "Thermotolerant"))),
           shape = 21,
           color = 'white',
           size = 4) +
  geom_point(aes(x = reorder(`Colony`, -combined),
                 y = mig_norm),
             size = 1,
             alpha = 0.15)  +
  facet_grid(cols = vars(fct_relevel(Species, c("<i>A. digitifera</i>",
                                        "<i>A. millepora</i>",
                                        "<i>A. </i>cf. <i>tenuis</i>"))),
             scales = 'free',
             space = 'free',
             axes = "all") +
  labs(y = 'Min-max normalized values',
       fill = "Group") +
  scale_fill_manual(values = Z_color_htol_int) +
  scale_y_continuous(limits = c(0,2.1), 
                     expand = expansion(0, 0)) +
  theme_classic() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        axis.line = element_line(),
        strip.text = element_markdown(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.background = element_rect(fill = 'transparent'),
        legend.key = element_blank(),
        plot.background = element_rect(fill = 'transparent'))
```

Export.

```{r}
#| eval: false
ggsave(plot = C_mor.mig_p,
       filename = '../Results/Figures/Figure 01C.png',
       dpi = 900,
       width = 11500, 
       height = 3000, 
       units = "px")
```
