---
title: "Rarefaction"
format: html
theme: 
  light: zephyr
  dark: cyborg
toc: true
page-layout: full
number-sections: true
number-depth: 3
editor: visual
self-contained: true
---

# Thermotolerance correlation with Symbiodiniaceae

# Notes

A naming convention for dataframes was made.

1.  Objects were named with a prefix per section arranged alphabetically.
2.  A 2nd prefix was added to arrange all the data accordingly:
    -   Dataframes: .**1**
    -   Plots: .**2**
    -   Stats: .**3** (those in dataframe format)
    -   Others:
        -   Vectors: .**vec**
        -   Trees: .**tre**
        -   Distances: .**dis**
        -   Functions: .**fun**
3.  If a dataframe will be used multiple times in case a raw dataframe is needed and an edited one should not supersede the raw one, add a suffix at the end (i.e., **\_1** for first edit, **\_2** for second....).
4.  Limit 3rd prefix names to three letters and if 2 prefixes are needed, separate with "**.**".
5.  If an object is just an auxiliary object, put a **Z** prefix.

```{r}
pacman::p_load(tidyverse,           
               janitor,
               cowplot,
               vegan,
               pairwiseAdonis,
               ggtext,
               rstatix,
               ggbeeswarm,
               ggtree,
               ape)
```

# Load prior work

```{r}
load("Rarefaction.RData")
```

# Load metadata

```{r}
#| warning: false


A.1_mta.htl <- 
  read_csv('../data/STab11.12.13_MinMaxValues.csv') |> 
  rename("Colony ID" = Colony)
```

## Make a color scheme

```{r}
Z.vec_col.spc <- 
  c("A. digitifera"='#221b1b',
    "A. millepora"="#AA4c0a",
    "A. tenuis"="#efb83a")
Z.vec_col.htl <- c("Thermotolerant" = "#B2182B","Intermediate" = "gray","Thermosensitive" = "#2166AC")
Z.vec_col.prf <- c("#203D3B","#6C9493","#F6E272","#F69312")
Z.vec_col.dur <- c("#B2182B","#ececec")
```

## Clean metadata

Among the Acropora species, only *A. tenuis* showed mortality in the control tanks during exposure. These will not be included in further analyses, cutting the A. tenuis colonies that will be included to 25.

These colonies are: ATEN07, ATEN08, ATEN09, ATEN24, ATEN25.

```{r}
Z.vec_mor.unr <- c("ATEN07","ATEN08","ATEN09","ATEN24","ATEN25")
```

# Load Symportal results

## ITS2 seqs

```{r}
#| warning: false
A.1_mta.its <-   
  read_tsv('../data/Fig03_ITS2.txt') |> 
  drop_na()
```

# Rarefaction

Convert ITS2 seqs to matrix.

```{r}
B.1_mta.its <- 
  A.1_mta.its |> 
  select(2,40:last_col()) |> 
  type_convert() |> 
  column_to_rownames("sample_name")

B.mat_its <- 
  B.1_mta.its |> 
  as.matrix()
```

Tidy.

```{r}
B.3_rare <- 
  rarecurve(B.mat_its, step = 100)

B.1_rare <- 
  B.3_rare |> 
  map_dfr(bind_rows) |> 
  bind_cols("Colony ID" = rownames(B.mat_its)) |> 
  filter(!(`Colony ID`) %in% Z.vec_mor.unr) |> 
  mutate(`Colony ID` = str_replace(`Colony ID`, 'ADIG', 'D'),
         `Colony ID` = str_replace(`Colony ID`, 'AMIL', 'M'),
         `Colony ID` = str_replace(`Colony ID`, 'ATEN', 'T')) |> 
  pivot_longer(!`Colony ID`) |> 
  drop_na() |> 
  mutate(n_seqs = as.numeric((str_replace(name, "N", "")))) |> 
  select(-name) |> 
  left_join(select(A.1_mta.htl, 1,2))
```

Plot.

```{r}
B.2_rare <- 
  B.1_rare |> 
  ggplot(aes(x = n_seqs,
             y = value,
             group = `Colony ID`,
             color = Species)) +
  geom_line(linewidth = 0.6,
            alpha = 0.5) +
  labs(y = "ITS2 richness",
       x = "Number of sequences") +
  theme_minimal() +
  scale_color_manual(values = Z.vec_col.spc) +
  theme(panel.grid = element_blank(),
        axis.line = element_line(),
        axis.ticks = element_line())
```

Save.

```{r}
ggsave(B.2_rare,
       filename = '../Results/Figures/2025/Rarefaction curves.png',
       dpi = 900,
       width = 12000, 
       height = 3000, 
       units = "px")
```

# Save work

```{r}
save.image("Rarefaction.RData")
```
