---
title: "Comparison of thermotolerance parameters"
format: html
theme: 
  light: zephyr
  dark: cyborg
toc: true
page-layout: full
number-sections: true
number-depth: 3
editor: visual
self-contained: true
---

# Notes

A naming convention for dataframes was made.

1.  Objects were named with a prefix per section arranged alphabetically.
2.  A 2nd prefix was added to arrange all the data accordingly:
    -   Dataframes: .**1**
    -   Plots: .**2**
    -   Stats: .**3** (those in dataframe format)
    -   Others:
        -   Vectors: .**vec**
        -   Trees: .**tre**
        -   Distances: .**dis**
        -   Functions: .**fun**
3.  If a dataframe will be used multiple times in case a raw dataframe is needed and an edited one should not supersede the raw one, add a suffix at the end (i.e., **\_1** for first edit, **\_2** for second....).
4.  Limit 3rd prefix names to three letters and if 2 prefixes are needed, separate with "**.**".
5.  If an object is just an auxiliary object, put a **Z** prefix.

# Load pkgs

```{r}
pacman::p_load(tidyverse,
               ggtext,
               ggbeeswarm,
               rstatix)
```

# Load prior work

```{r}
load("Comparison of thermotolerance parameters.RData")
```

# Load data

```{r}
A.1_thermo.params <- 
  read_csv("../data/STab03_ThermotoleranceAll.csv") |> 
  rename("Colony ID" = Colony) |> 
  select(1:4) |> 
  rename("IOM" = 3,
         "CC" = 4)

A.1_IOMD10 <- 
  read_csv("../data/STab03_IOM.csv") |> 
  select(5,1,10)|> 
  mutate(Colony = str_replace(Colony, 'ADIG', 'D'),
         Colony = str_replace(Colony, 'AMIL', 'M'),
         Colony = str_replace(Colony, 'ATEN', 'T')) |> 
  rename("Colony ID" = Colony,
         "IOMD10" = 3)

A.1_thermo.params_1 <- 
  A.1_thermo.params |> 
  left_join(A.1_IOMD10)
```

Tidy.

```{r}
Z.fun_minmax <- 
  function(x) {
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
}

Z.vec_mor.unr <- c("T07","T08","T09","T24","T25")

B.1_thermo.params <- 
  A.1_thermo.params_1 |> 
  filter(!`Colony ID` %in% Z.vec_mor.unr) |> 
  mutate(mmIOM = Z.fun_minmax(IOMD10),
         mmCC = Z.fun_minmax(CC)) |> 
  rowwise() |> 
  mutate(mmCombined = mmIOM + mmCC) |> 
  ungroup()
```

# Plot

Make a list of colors.

```{r}
Z.vec_col.sp <- c("Adig"='#221b1b',
                   "Amil"="#AA4c0a",
                   "Aten"="#efb83a")
```

Among the Acropora species, only *A. tenuis* showed mortality in the control tanks during exposure. These will not be included in further analyses, cutting the A. tenuis colonies that will be included to 25.

These colonies are: ATEN07, ATEN08, ATEN09, ATEN24, ATEN25.

```{r}
C.1_thermo.params <- 
  B.1_thermo.params |> 
  select(!c(3:5)) |> 
  pivot_longer(!c(1:2),
               names_to = "ThermotoleranceParameter",
               values_to = "ThermotoleranceMeasure")
```

Save.

```{r}
C.1_thermo.params |> 
  write_csv("../Results/Tables/2025/Thermotolerance parameters in all colonies - Regardless of species.csv")
```

Plot.

```{r}
C.2_thermo.params <- 
  C.1_thermo.params |> 
  mutate(Species = case_when(Species == "A. tenuis" ~ "Aten",
                             Species == "A. millepora" ~ "Amil",
                             .default = "Adig")) |> 
  ggplot(aes(x = fct_relevel(Species, c("Adig",
                                        "Amil",
                                        "Aten")),
             group = Species,
             y = ThermotoleranceMeasure,
             fill = Species,
             color = Species)) +
  geom_quasirandom(method = 'pseudorandom',
                   alpha = 0.25,
                   size = 2, 
                   width = 0.25) +
  facet_wrap(~ fct_relevel(ThermotoleranceParameter, c("mmIOM", "mmCC", "mmCombined")),
             scale = "free",
             axes = "all") +
  stat_summary(fun = 'median',
               size = 1,
               shape = 21,
               color = 'white',
               stroke = 0.75) +
  ylab(label = 'Min-max normalized values') +
  scale_fill_manual(values = Z.vec_col.sp) +
  scale_color_manual(values = Z.vec_col.sp) +
  theme_classic() +
  theme(panel.grid = element_blank(),
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        axis.line = element_line(),
        axis.text.x = element_markdown(),
        axis.ticks.x = element_blank(),
        legend.background = element_rect(fill = 'transparent'),
        legend.key = element_blank(),
        plot.background = element_rect(fill = 'transparent', color = "transparent"),
        legend.title = element_blank(),
        legend.text = element_markdown(),
        legend.position = "none")
```

Export.

```{r}
ggsave(C.2_thermo.params,
       filename = '../Results/Figures/2025/Comparison of thermotolerance parameters.png',
       dpi = 900,
       width = 4250, 
       height = 2200, 
       units = "px")
```

# Stats

Kruskal-Wallis.

```{r}
D.3_kruskal <- 
  C.1_thermo.params |> 
  group_by(ThermotoleranceParameter) |> 
  kruskal_test(ThermotoleranceMeasure ~ Species)
```

Dunn's test.

```{r}
E.3_dunns <- 
  C.1_thermo.params |> 
  group_by(ThermotoleranceParameter) |> 
  dunn_test(ThermotoleranceMeasure ~ Species,
            p.adjust.method = 'BH')
```

Median and IQR.

```{r}
F.1_median.iqr <- 
  C.1_thermo.params |> 
  group_by(Species, ThermotoleranceParameter) |> 
  summarise(Q1 = round(quantile(ThermotoleranceMeasure, 0.25),2),
            Q3 = round(quantile(ThermotoleranceMeasure, 0.75),2),
            Median = median(ThermotoleranceMeasure)) |> 
  mutate(IQR_range = paste0(Q1, " - ", Q3))
```

Export.

```{r}
#| eval: false
write_csv(E.3_dunns,
          '../Results/Tables/2025/Dunns test.csv')
```

## Summary stats

```{r}
G.1_sum.stats <- 
  C.1_thermo.params |> 
  group_by(Species, ThermotoleranceParameter) |> 
  summarise(Q1 = round(quantile(ThermotoleranceMeasure, 0.25),2),
            Q3 = round(quantile(ThermotoleranceMeasure, 0.75),2),
            Median = median(ThermotoleranceMeasure)) |> 
  mutate(IQR_range = paste0(Q1, " - ", Q3))
```

Export.

```{r}
G.1_sum.stats |> 
  write_csv('../Results/Tables/2025/Thermotolerance parameters - Summary stats.csv')
```

# Save work

```{r}
save.image("Comparison of thermotolerance parameters.RData")
```
